\TUDoption{abstract}{section, multiple}

% TODO: Replace the placeholders with your actual abstracts

\newcommand*{\Abstreng}{
  High-resource languages benefit from abundant data, allowing named entity recognition tasks to achieve
  high accuracy. However, low-resource languages and specialized domains face challenges due to
  limited training data. Current multilingual capabilities of models are insufficient for
  low-resource languages, prompting ongoing research. Approaches to address NER in these contexts
  are split into model-based and data-based, with projection-based methods being a significant subset
  of the latter. These methods involve translating sentences into high-resource languages, labeling
  them, and projecting labels back to the original. Despite many proposals for projection, this step
  hasn't been explicitly formalized as an optimization problem, often relying on heuristics. This study
  addresses this gap by formulating the projection step as an integer linear programming problem within
  the cross-lingual NER pipeline. This ILP approach optimizes matching of source to target entities
  using varied cost strategies: alignment-based, NER-model-based, and translation-based
  scores,— integrated to mitigate individual limitations. While the ILP problem's complexity remains
  an open question, a proposed greedy algorithm performs effectively in practice. Evaluations using
  Europarl-NER and MasakhaNER2 datasets show our ILP method excels over heuristic and model transfer
  techniques, especially with fused scores.
}

% \newcommand*{\Abstrger}{}

\begin{abstract}[pagestyle=empty.tudheadings]
  \Abstreng
  % \nextabstract[ngerman]
  % \Abstrger

\end{abstract}
