\chapter{Appendix I}
\label{sec:appendix}

This appendix presents the overall runtimes, measured in seconds, for all
experiments conducted on the Europarl-NER dataset.
All results were obtained on the HPC system of TU Dresden, specifically
on the "Alpha Centauri" partition, where each job
utilized 1 NVIDIA A100-SXM4 GPU, 64 GB of RAM, and 8 cores of AMD EPYC CPU 7352.

\begin{table}[ht]
  \centering
  \input{mainmatter/tables/europarl_heuristic_runtime}
  \caption{Overall runtime is seconds for word-to-word alignments-based heuristic
  algorithm with different hyperparameter  on the Europarl NER dataset}
  \label{tab:europarl_heur_runtime}
\end{table}

It should be noted that all word-to-word alignments were computed
and saved prior to executing any part of the pipeline; consequently,
the time required for their computation is not reflected in the
tables. As a result, the comparison between the alignment-based pipeline
and both the NER and NMT may not accurately represent their
relative runtimes. However, it is generally anticipated that
the computation of alignments will demand a duration comparable
to that of executing a NER model; thus, the runtimes for the
NER and alignment-based pipelines are expected to be similar.

Additionally, it is worth to point out, that according to Theorem \ref{thm:reduce_T_for_align},
the set of candidates \( T \) has been reduced for all ILP-based experiments that
utilize only word-to-word alignments, resulting in reduced runtimes. Furthermore, the initial
computation of NMT scores for the dataset (which corresponds to the \textit{nmt}
pipeline with constraints of type \( \leq 1 \) and the GUROBI solver) caches certain
precomputations. As a result, all subsequent runs involving NMT scores require
less time.

\begin{table}[t]
  \centering
  \input{mainmatter/tables/europarl_ilp_runtime}
  \caption{Overall runtime is seconds for the ILP based projection pipelines
  on the Europarl NER dataset}
  \label{tab:europarl_ilp_runtime}
\end{table}
