\chapter{Methodology}
\label{sec:methodology}
The overall idea of formalization the projection step of the XLNER pipeline
is to represent it as a matching the source entities and ranges of words
of the target sentence, called candidates. It is shown on the figure \ref{fig:cand_matching}.
The set of source entities that we need to project is given, so let's take some set of
continuous word ranges from the original sentence in the target language
and consider them as a possible projection of some entities. Then compute
quantities how likely a source entity should be projected onto every candidate.
And then we should take such a combination of source entities-target candidates
that maximize overall sum of selected tuples. But we need to take in mind
natural constraints that should be ensured. Assume two source entities
projects onto overlapping candidates, then we cannot precisely answer which source
entity overlapped words are projected to, therefore we should prohibit projection
onto overlapped candidates. Also, it makes sense to limit how many candidates every
source entity can be projected to, since usually we expect 1-to-1 correspondence
between source entities and entities of the target sentence. Based on this simple
properties let's build an integer linear optimization problem and study it.

\begin{figure*}[ht]
  \centering
  \begin{tikzpicture}[node distance=-0.1,
      every node/.style={text centered,
        text height=2ex,
        text depth=.25ex,
      },
      loc/.style={fill=orange!30, rounded rectangle, label={[anchor=center,font=\tiny\bfseries\sffamily]above:#1-LOC}},
      per/.style={fill=green!30, rounded rectangle, label={[anchor=center,font=\tiny\bfseries\sffamily]above:#1-PER}},
    cand/.style={fill=blue!30, rounded rectangle},]

    \node[per={B}, rounded rectangle east arc=none](George_src){George};
    \node[per={I}, rounded rectangle west arc=none, right=of George_src](Washington_src){Washington};
    \node[right=of Washington_src](is_src){is};
    \node[right=of is_src](the_src){the};
    \node[right=of the_src](first_src){first};
    \node[right=of first_src](president_src){president};
    \node[right=of president_src](of_src){of};
    \node[right=of of_src](the_src){the};
    \node[loc={B}, rounded rectangle east arc=none, right=of the_src](United_src){United};
    \node[loc={I}, rounded rectangle west arc=none, right=of United_src](States_src){States};

    \node[cand, rounded rectangle east arc=none, below=of George_src, yshift=-1.5cm](George_tgt){George};
    \node[cand, rounded rectangle west arc=none, right=of George_tgt](Washington_tgt){Washington};
    \node[right=of Washington_tgt](ist_tgt){ist};
    \node[right=of ist_tgt](der_tgt){die};
    \node[right=of der_tgt](erste_tgt){erste};
    \node[right=of erste_tgt](president_tgt){Präsident};
    \node[right=of president_tgt](der_tgt){der};
    \node[cand, rounded rectangle east arc=none, right=of der_tgt](Vereinigten_tgt){Vereinigten};
    \node[cand, rounded rectangle west arc=none, right=of Vereinigten_tgt](Staaten_tgt){Staaten};

    \node[text=gray, font=\scriptsize, above=of first_src, yshift=0.2cm, xshift=0.2cm](source){Source labeled sentence};
    \node[text=gray, font=\scriptsize, below=of source, yshift=-3cm]{Original sentence with extracted candidates};

    \draw[->] (George_src.south east) -- node[left]{\(c_{11}\)} (George_tgt.north east);
    \draw[->] (United_src.south east) -- node[right]{\(c_{22}\)} (Vereinigten_tgt.north east);
    \draw[->] (George_src.south east) -- node[above left, yshift=0.1cm, xshift=-0.2cm]{\(c_{12}\)} (Vereinigten_tgt.north east);
    \draw[->] (United_src.south east) -- node[below left]{\(c_{21}\)} (George_tgt.north east);
  \end{tikzpicture}
  \caption{Illustration of the proposed idea of matching source entities and candidates in the target sentence}
  \label{fig:cand_matching}
\end{figure*}

\section{Formulation of the ILP problem}
First of all let's start with a definition of overlapping.

\begin{definition}[Relation of overlapping] \label{def:overlapping}
  We consider that target candidates \linebreak
  \( \tgt{p_1} = (i_{\tgt{p_1}}, j_{\tgt{p_1}}), \tgt{p_2} = (i_{\tgt{p_2}}, j_{\tgt{p_2}}) \in T \) overlap if and only if
  \( i_{\tgt{p_1}} \leq j_{\tgt{2}} \land i_{\tgt{p_2}} \leq j_{\tgt{1}} \), i. e. sets of word indices are
  not disjoint. We will denote that candidates overlap using
  the following notation: \( \tgt{p_1} \cap \tgt{p_2} \neq \emptyset \).
\end{definition}

Having defined all necessary objects we can finally formulate the ILP problem for a
projection step of the XLNER pipeline that aligns with our requirements.

Let \( S \) be a set of source entities, \( T \) -- set of target candidates and \( \cap \subset T^2 \) is
a relation of overlapping, then the projection ILP problem is the following:
\begin{align}
  \label{eq:objective}
  & \max\limits_x \sum\limits_{(\src{p}, \tgt{p}) \in S \times T} c_{\src{p}, \tgt{p}} x_{\src{p}, \tgt{p}}                                             \\
  & \text{subject to} \notag                                                                                                                            \\
  \label{eq:num_proj_const}
  & \sum\limits_{\tgt{p} \in T} x_{\src{p}, \tgt{p}} \lessgtr n_{proj}                                      & \forall \src{p} \in S                     \\
  \label{eq:non_overlap_const}
  & x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1
  & \forall (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T)                                                                            \\
  \label{eq:binary_prog}
  & x_{\src{p}, \tgt{p}} \in \{ 0, 1 \}                                                                     & \forall (\src{p}, \tgt{p}) \in S \times T
\end{align}
where \( \hat{\Pi}(S, T) \) is a set of source entities -- target candidates combinations that can not be
projected onto together since they are overlapping. This set is defined as follows:
\begin{equation}  \label{eq:overlapping_set}
  \begin{aligned}
    \hat{\Pi}(S, T) = \Big\{ (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \Big| \src{p_1}, \src{p_2} \in S, \tgt{p_1}, \tgt{p_2} \in T, \quad \tgt{p_1} \cap \tgt{p_2} \neq \emptyset, \\
    (\src{p_1} \neq \src{p_2}) \lor (\tgt{p_1} \neq \tgt{p_2}) \Big\}
  \end{aligned}
\end{equation}

Here every variable \( x_{\src{p}, \tgt{p}} \) represents whether the source entity \( \src{p} \in S \)
is projected to the target candidates \( \tgt{p} \in T \).
The set of constraints \eqref{eq:non_overlap_const} ensures that it is impossible to have a situation when
two or one source entity are being projected onto overlapping target candidates.
And other set of constraints \eqref{eq:num_proj_const} limits number of projections
for every source entity. Note, that the inequality sign is not fixed since any form
could make sense: less, less or equal and equal can be used to ensure that we don't
project a source entity to as much available candidates as possible as long as it
increase the objection value, and great or great or equal can be a way to enforce
that a source entiry will be projected at least specified number of times, e.g. once,
and solver will not simply ignore it. It is clear that \( n_{proj} = 0 \) is a corner
case and usually doesn't make sense except greater inequality when it simply
eliminates any limits and equivalent to not have any constraints of such type.

Also it is worth to note that definition of the ILP problem is not bounded to a specific form of sets \( S, T \)
and overlapping relation \( \cap \subset T^2 \), so given any sets and overlapping relation with the same properties
as for \ref{def:overlapping}, the ILP problem won't change since it doesn't use explicitly word indices.

However the number of constraints \eqref{eq:non_overlap_const} scales as a cube with respect to
length of the target sentence, assuming that we use all possible continuous subranges as candidates, and as
a square with respect to the number of source entities.
It can make this problem harder to solve. Therefore we need to try to lower number of constraints
of this type. For this let's take a look at the overlapping relation that a basis of it.
\begin{lemma} \label{lemma:not_transitive}
  The relation of overlapping is not transitive.
\end{lemma}
\begin{proof}
  We will proof this by showing a counterexample.
  Assume the following target candidates
  \[
    a = (2, 4) \quad b = (1, 2) \quad c = (4, 5)
  \]
  By the definition of the overlapping relation \( a \cap b \neq \emptyset \) and
  \( a \cap c \neq \emptyset \) but \( b \) and \( c \) are not overlapping.
\end{proof}

This fact leads to the impossibility to split all target candidates into groups of mutually
overlapped candidates and choice at maximum one candidate from every group to match with a source entity.
\begin{corollary}
  It is impossible to partition set of target candidates such that every partition consists of
  candidates that are pairwise overlapping with each other and there are no overlapping candidates
  that are in different partitions.
\end{corollary}
\begin{proof}
  By Lemma \ref{lemma:not_transitive} the overlapping relation is not transitive and therefore
  is not an equivalence relation that implies we can not construct a quotient set.
\end{proof}

Nevertheless it is possible to make number of constraints \eqref{eq:non_overlap_const} lower.
The idea behind the reduction is that if it is impossible to project one or any two source entities to
overlapped candidates it means that it impossible to project any number of source entities to these candidates.
Thus we can just sum up constraints over all source entities.
We can do it because all variables \( x_{\src{p}, \tgt{p}} \) are binary and therefore
can not be negative.
\begin{theorem}
  Set of constraints \eqref{eq:non_overlap_const} is satisfied if and only if
  the following sets of reduced constraints are satisfied:
  \begin{align*}
    & \sum\limits_{\src{p} \in S} (x_{\src{p}, \tgt{p_1}} + x_{\src{p}, \tgt{p_2}}) \leq 1                                & \forall (\tgt{p_1}, \tgt{p_2}) \in \Pi(T) \\
    & \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} \leq 1
    & \forall \tgt{p} \in T \Big| \nexists \tgt{p_2} \in T: \tgt{p} \neq \tgt{p_2}, \tgt{p} \cap \tgt{p_2} \neq \emptyset                                             \\
  \end{align*}
  where
  \begin{equation*}
    \Pi(T) = \left\{ (\tgt{p_1}, \tgt{p_2}) \Big| \tgt{p_1}, \tgt{p_2} \in T,
      \quad \tgt{p_1} \cap \tgt{p_2} \neq \emptyset,
    \tgt{p_1} \neq \tgt{p_2} \right\}
  \end{equation*}
\end{theorem}
\begin{proof}
  \textit{Necessity:} Assume that constraints \eqref{eq:non_overlap_const} are satisfied, but
  the proposed constraints are not. Then there are two options:
  \[
    \exists (\tgt{p_1}, \tgt{p_2}) \in \Pi(T) \Bigg| \quad
    \sum\limits_{\src{p} \in S} (x_{\src{p}, \tgt{p_1}} + x_{\src{p}, \tgt{p_2}}) > 1
  \]
  or
  \[
    \exists \tgt{p} \in T \Big|
    \nexists \tgt{p_2} \in T: \tgt{p} \neq \tgt{p_2}, \tgt{p} \cap \tgt{p_2} \neq \emptyset
    \Bigg| \quad \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} > 1
  \]

  Consider the first option:
  \begin{align*}
    & \sum\limits_{\src{p} \in S} (x_{\src{p}, \tgt{p_1}} + x_{\src{p}, \tgt{p_2}}) > 1
    \stackrel{\eqref{eq:binary_prog}}{\implies}                                                                     \\
    & \exists \src{p_1}, \src{p_2} \in S \Big| x_{\src{p_1}, \tgt{p_1}} = 1, x_{\src{p_2}, \tgt{p_2}} = 1 \implies \\
    & x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} = 2 > 1
  \end{align*}
  that contradicts our assumption that constraints \eqref{eq:non_overlap_const} are satisfied.

  The same result we can obtain for the second case:
  \begin{align*}
    & \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} > 1
    \stackrel{\eqref{eq:binary_prog}}{\implies}                                                                                           \\
    & \exists \src{p_1}, \src{p_2} \in S, \src{p_1} \neq \src{p_2} \Big| x_{\src{p_1}, \tgt{p}} = 1, x_{\src{p_2}, \tgt{p}} = 1 \implies \\
    & x_{\src{p_1}, \tgt{p}} + x_{\src{p_2}, \tgt{p}} = 2 > 1
  \end{align*}

  Thus we proved necessity.

  \textit{Sufficiency:}
  Suppose that the following constraints are satisfied:
  \[
    \sum\limits_{\src{p} \in S} (x_{\src{p}, \tgt{p_1}} + x_{\src{p}, \tgt{p_2}}) \leq 1
    \qquad \forall (\tgt{p_1}, \tgt{p_2}) \in \Pi(T)
  \]
  Then the sum could be equal to \( 0 \) or \( 1 \). If it is zero then none entities are projected onto
  any of these overlapped candidates and therefore constraints \eqref{eq:non_overlap_const} are not violated.
  Let's look at case when the sum is equal to \( 1 \).
  \begin{equation*} \label{eq:derivation_of_const_reduction}
    \begin{aligned}
      & \sum\limits_{\src{p} \in S} (x_{\src{p}, \tgt{p_1}} + x_{\src{p}, \tgt{p_2}}) = 1
      \stackrel{\eqref{eq:binary_prog}}{\implies}                                                       \\
      & \exists! \src{p} \in S, \exists! \tgt{p} \in \{ \tgt{p_1}, \tgt{p_2}\}
      \Bigg| x_{\src{p}, \tgt{p}} = 1 \implies                                                          \\
      & \forall \src{p_1}, \src{p_2} \in S, x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1
    \end{aligned}
  \end{equation*}
  Therefore we proved that
  \[
    x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1 \qquad
    \forall (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi'}(S, T)
  \]
  where
  \begin{align*}
    \hat{\Pi'}(S, T) = \Big\{ (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \Big| \src{p_1}, \src{p_2} \in S, \tgt{p_1}, \tgt{p_2} \in T, \quad \tgt{p_1} \cap \tgt{p_2} \neq \emptyset, \\
    \tgt{p_1} \neq \tgt{p_2} \Big\}
  \end{align*}

  But it is not the same as set \eqref{eq:overlapping_set}:
  \[
    \hat{\Pi}(S, T) \setminus \hat{\Pi'}(S, T) = \Big\{ (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \Big| \src{p_1}, \src{p_2} \in S, \tgt{p_1}, \tgt{p_2} \in T, \tgt{p_1} = \tgt{p_2},
    \src{p_1} \neq \src{p_2} \Big\}
  \]
  So we need to check whether constraints \eqref{eq:non_overlap_const} are satisfied on this set difference.

  First of all, let's notice that
  \begin{multline*}
    \forall (\tgt{p_1}, \tgt{p_2}) \in \Pi(T)
    \qquad
    \sum\limits_{\src{p} \in S} (x_{\src{p}, \tgt{p_1}} + x_{\src{p}, \tgt{p_2}})
    =
    \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p_1}} +
    \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p_2}} \leq 1
    \implies                        \\
    \forall \tgt{p} \in T \Bigg| \exists \tgt{p_2} \in T: \tgt{p} \neq \tgt{p_2}, \tgt{p} \cap \tgt{p_2} \neq \emptyset
    \quad
    \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} \leq 1
  \end{multline*}
  Therefore there exists at most one source entity \( \src{p} \in S \) that are projected
  to the target candidate \( \tgt{p} \).
  And using exactly the same derivation as above we get:
  \begin{multline*}
    \forall \tgt{p} \in T \Bigg| \exists \tgt{p_2} \in T: \tgt{p} \neq \tgt{p_2}, \tgt{p} \cap \tgt{p_2} \neq \emptyset
    \quad
    \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} \leq 1
    \stackrel{\eqref{eq:binary_prog}}{\implies}                                                                                                                                                       \\
    \begin{aligned}
      & \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} = 0
      \stackrel{\eqref{eq:binary_prog}}{\implies}
      \forall \src{p} \in S, x_{\src{p}, \tgt{p}} = 0         \\
      \text{or}                                               \\
      & \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} = 1
      \stackrel{\eqref{eq:binary_prog}}{\implies}
      \exists! \src{p} \in S, x_{\src{p}, \tgt{p}} = 1        \\
    \end{aligned} \implies \\
    x_{\src{p_1}, \tgt{p}} + x_{\src{p_2}, \tgt{p}} \leq 1 \qquad
    \begin{aligned}
      & \forall \src{p_1}, \src{p_1} \in S, \src{p_1} \neq \src{p_2}                                                        \\
      & \forall \tgt{p} \in T \Bigg| \exists \tgt{p_2} \in T: \tgt{p} \neq \tgt{p_2}, \tgt{p} \cap \tgt{p_2} \neq \emptyset
    \end{aligned}
  \end{multline*}
  So, only the case, when entity has no overlapped entity that are
  not equal to it, remains. But this case is fully covered by the second part of the proposed reduced constraints:
  \[
    \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} \leq 1 \qquad
    \forall \tgt{p} \in T \Big| \nexists \tgt{p_2} \in T: \tgt{p} \neq \tgt{p_2}, \tgt{p} \cap \tgt{p_2} \neq \emptyset
  \]
  And by considering that this sum can be either \( 0 \) or \( 1 \) and making
  exactly the same derivation we conclude that it implies that
  constraints \eqref{eq:non_overlap_const} are satisfied.
\end{proof}

The theorem allows us to reduce number of constraints and make them scale as a constant
with respect to the number of source entities. As a result we end up with the following final formulation of the ILP problem for the projection step
of the XLNER pipeline.
\begin{equation} \label{eq:ilp}
  \begin{aligned}
    & \max\limits_x \sum\limits_{(\src{p}, \tgt{p}) \in S \times T} c_{\src{p}, \tgt{p}} x_{\src{p}, \tgt{p}}                                                                                                                       \\
    & \text{subject to}                                                                                                                                                                                                             \\
    & \sum\limits_{\tgt{p} \in T} x_{\src{p}, \tgt{p}} \lessgtr n_{proj}                                      & \forall \src{p} \in S                                                                                               \\
    & \sum\limits_{\src{p} \in S} (x_{\src{p}, \tgt{p_1}} + x_{\src{p}, \tgt{p_2}}) \leq 1                    & \forall (\tgt{p_1}, \tgt{p_2}) \in \Pi(T)                                                                           \\
    & \sum\limits_{\src{p} \in S} x_{\src{p}, \tgt{p}} \leq 1                                                 & \forall \tgt{p} \in T \Big| \nexists \tgt{p_2} \in T: \tgt{p} \neq \tgt{p_2}, \tgt{p} \cap \tgt{p_2} \neq \emptyset \\
    & x_{\src{p}, \tgt{p}} \in \{ 0, 1 \}                                                                     & \forall (\src{p}, \tgt{p}) \in S \times T
  \end{aligned}
\end{equation}

\section{Candidates extraction}
Whereas the set \( S \) of source entities are given, since all source entities have been
extracted on the previous step of the XLNER pipeline, the construction of the set \( T \)
of target candidates is an open question.

The simplest way for candidates extraction is to consider all possible n-grams
(continuous ranges of words) of the target sentence as candidates. This approach ensures
that there won't be any possibility that we don't include any actual target entity in the set
target candidates. But when it comes to the computational point of view it can be a problem since
number of all n-grams scales quadratically. And from the application point of view most of
the candidates generated in such way are not good nominee to be a real target entity.

One of the simplest idea to handle this problem is to limit the maximum possible length of the candidate.
From the point of NER it is natural to expect that real target entities usually do not exceed some
predefined number. For example it is unlikely that a person's name will contain more than 10 words or
if we have 3 source entities there is a real target entity that contains all words of the target sentence.

Thus, the algorithm for candidates extraction wilt bounded maximum length of n-grams are the following.
\begin{algorithm}
  \KwData{\( n \in \mathbb{N} \) -- number of words in the target sentence,
  \(M \leq n \in \mathbb{N} \) -- maximum length of a target candidate}
  \KwResult{\( T \) -- set of target candidates}

  \( T \gets \emptyset \) \;
  \For{\(i \gets 0 \) \KwTo \( n \)}{
    \( m \gets \min(s + M, n) \) \;
    \For{\(j \gets i \) \KwTo \( m \)}{
      \( T \gets T \cup \{ (i, j) \} \) \;
    }
  }
  \caption{Bounded length n-gram candidates extraction}
  \label{alg:ngram_extraction}
\end{algorithm}

An alternative approach for candidates extraction is to use some model to generate candidates.
TProjection uses fine-tuned T5 model with beam-search to generate candidates. In theory any LLM can be finetuned
to do it. But the main drawback of this approach is that we need to finetune these models for which we need
a labeled dataset in a target language or rely on cross-lingual model transfer.
But in this case it will be easier and more effective to train a model to label
target entities itself rather than generate candidates for the projection step of the
XLNER pipeline. In addition, autoregressive models are very resource consuming and it is
inefficient to use them only for candidates extraction.

Therefore it make sense also consider multilingual encoder-only Transformer model with a high recall,
but we still need to train these model to extract candidates. In the following chapters we will not consider all
these alternatives and leave for further works.

\section{Matching scores}
So far we have formulated and discussed the ILP problem \eqref{eq:ilp} that projects source entities onto
target candidates using a matching score \( c_{\src{p}, \tgt{p}} \). This score represents
how likely given source entity \( \src{p} \in S \) should be projected onto the target candidate \( \tgt{p} \in T \).
But the question how to compute all these scores remains unanswered. In this section we will propose
some options to evaluate them.

\subsection{Alignment-based score}
In the chapter \ref{sec:background} it has been showed that word-to-word alignments can be used
for the projection step of the XLNER pipeline. But it was done by the heuristic algorithm.
Nevertheless we can try incorporate them into the proposed ILP problem by calculating scores of matching with use of
word-to-word alignments.

The matching score that inspired from this idea is the following:
\begin{equation} \label{eq:align_cost}
  c_{\src{p}, \tgt{p}}^{align} =
  \frac{\sum\limits_{k=i_{\src{p}}}^{j_{\src{p}}} \sum\limits_{l=i_{\tgt{p}}}^{j_{\tgt{p}}} a_{kl}}
  {(j_{\src{p}} - i_{\src{p}}) + (j_{\tgt{p}} - i_{\tgt{p}})}
\end{equation}

The form of the score based on the natural properties that we can expect from score that based on word-to-word alignments.
First of all, the more aligned words between source entity and target candidates the higher should be a score.
Secondly, we can not divide number of aligned word by length of only source entity then
score will be higher for just longer target candidates.
And the third property is especially useful: if we consider two candidates where one is a
substring of other, but total number of aligned words between words of
source entity and these candidates are the same, we should prefer smaller candidate.
And the score \eqref{eq:align_cost} satisfies this requirement.
\begin{lemma} \label{lemma:align_cost_decrease}
  Suppose that for some pair of source entity \( \src{p} \in S \) and target candidate \( \tgt{p} \in T \)
  the score \eqref{eq:align_cost} equal to c, then if there is an extended by 1 word to left or
  right candidate \( \tgt{\hat{p}} \in T \) such as this word is not aligned to any word of source
  entity, the score \eqref{eq:align_cost} between source entity and this target candidate
  will be lower than \( c \).
\end{lemma}
\begin{proof}
  Consider extension to the right, then \( \tgt{\hat{p}} = (i_{\tgt{p}}, j_{\tgt{p}} + 1) \). Since it is
  given that this additional word is not aligned to any word of the source entity it means that:
  \[
    \sum\limits_{k=i_{\src{p}}}^{j_{\src{p}}} a_{k,j_{\tgt{p}} + 1} = 0
  \]
  Then
  \begin{multline*}
    c_{\src{p}, \tgt{\hat{p}}}^{align} =
    \frac{\sum\limits_{k=i_{\src{p}}}^{j_{\src{p}}} \sum\limits_{l=i_{\tgt{p}}}^{j_{\tgt{p}} + 1} a_{kl}}
    {(j_{\src{p}} - i_{\src{p}}) + (j_{\tgt{\hat{p}}} - i_{\tgt{\hat{p}}})} =                                \\
    \frac{\sum\limits_{k=i_{\src{p}}}^{j_{\src{p}}} \sum\limits_{l=i_{\tgt{p}}}^{j_{\tgt{p}}} a_{kl}}
    {(j_{\src{p}} - i_{\src{p}}) + (j_{\tgt{p}} + 1 - i_{\tgt{p}})} +
    \frac{\sum\limits_{k=i_{\src{p}}}^{j_{\src{p}}} a_{k,j_{\tgt{p}} + 1}}
    {(j_{\src{p}} - i_{\src{p}}) + (j_{\tgt{p}} + 1 - i_{\tgt{p}})} =                                        \\
    \frac{\sum\limits_{k=i_{\src{p}}}^{j_{\src{p}}} \sum\limits_{l=i_{\tgt{p}}}^{j_{\tgt{p}}} a_{kl}}
    {(j_{\src{p}} - i_{\src{p}}) + (j_{\tgt{p}} - i_{\tgt{p}}) + 1} <
    \frac{\sum\limits_{k=i_{\src{p}}}^{j_{\src{p}}} \sum\limits_{l=i_{\tgt{p}}}^{j_{\tgt{p}}} a_{kl}}
    {(j_{\src{p}} - i_{\src{p}}) + (j_{\tgt{p}} - i_{\tgt{p}})} = c
  \end{multline*}
  The derivation for the extension to the left where \( \tgt{\hat{p}} = (i_{\tgt{p}} - 1, j_{\tgt{p}}) \) is structurally the same, so we
  proved the lemma.
\end{proof}

\begin{corollary} \label{col:shrink_cand}
  The cost of the extended candidate \( \tgt{p_{+n}} \) of the target candidate \( \tgt{p} \) to the left or right with
  \( n \in \mathbb{N} \) non-aligned to any word of the source entity \( \src{p} \in S \) words is lower than
  the cost of the original candidate.
\end{corollary}
\begin{proof} Let's prove it by the mathematical induction for the case of extension to the right (for the left it will the same).

  \textit{Base:} By the lemma \ref{lemma:align_cost_decrease} \( c_{\src{p}, \tgt{p}}^{align} > c_{\src{p}, \tgt{p_{+1}}}^{align} \).

  \textit{Induction step:} Assume that for some \( k \in \mathbb{N} \) holds \( c_{\src{p}, \tgt{p}}^{align} > c_{\src{p}, \tgt{p_{+k}}}^{align} \).
  Then by the lemma~\ref{lemma:align_cost_decrease} \( c_{\src{p}, \tgt{p_{+k}}}^{align} > c_{\src{p}, \tgt{p_{+(k+1)}}}^{align} \) and therefore
  \( c_{\src{p}, \tgt{p}}^{align} > c_{\src{p}, \tgt{p_{+(k+1)}}}^{align} \).

  Thus, by the mathematical induction, \( c_{\src{p}, \tgt{p}}^{align} > c_{\src{p}, \tgt{p_{+n}}}^{align}   \)

\end{proof}

The main advantages of this property is that it is possible to shrink the set of target candidates without losing
optimal solutions: we can consider as candidates only subranges that are between the leftmost and rightmost words that are aligned to any
word of any source entity.
\begin{theorem}
  Let \( m, M \in \mathbb{N} \) be indices of a leftmost and rightmost aligned word to any source entity, then there is
  an optimal solution for the problem \eqref{eq:ilp} where constraints \eqref{eq:num_proj_const} has a form
  of \( < \) or \( \leq \), such that it projects source entity to a target candidate from the set \( T \) generated by the algorithm
  \ref{alg:ngram_extraction} where any word's index is inside the range \( [m, M] \).
\end{theorem}
\begin{proof}
  Suppose that there is no such an optimal solution that satisfies this requirement. It means that there exists
  a source entity \( \src{p} \in S \) and a source candidate
  \( \tgt{\hat{p}} \in S \) such that \( (i_{\tgt{\hat{p}}} < m) \lor (M < j_{\tgt{\hat{p}} < m}) \)
  and \( x_{\src{p}, \tgt{\hat{p}}} = 1 \) for an optimal solution \( x \).

  This case can be divided into two parts: when \( \tgt{\hat{p}} \) has no words that aligned to words of source entities or
  is and vice versa.

  Consider the first option, then since \( \tgt{\hat{p}}  \) consists of words that are not aligned to any word
  of source entities the alignment-based matching cost according to the definition \eqref{eq:align_cost} equals to \( 0 \).
  Then if we take such a solution \( x^* \) that it is equal to this optimal solution \( x \) everywhere except
  for the target candidate \( \tgt{\hat{p}}  \), i. e. \( x_{\src{p}, \tgt{\hat{p}}} = 0 \). Thus the objective function
  remains the same because cost of matching is zero. And we don't violate non-overlapping constraints~\eqref{eq:non_overlap_const}:
  \begin{align*}
    & \forall \src{p_1} \in S, \tgt{p_1} \in T \Big| (\src{p_1}, \src{p}, \tgt{p_1}, \tgt{\hat{p}}) \in \hat{\Pi}(S, T) \\
    & x_{\src{p_1}, \tgt{p_1}} + x_{\src{p}, \tgt{\hat{p}}} =
    x_{\src{p_1}, \tgt{p_1}} + 1 \leq 1 \implies                                                                         \\
    & x^*_{\src{p_1}, \tgt{p_1}} + x^*_{\src{p}, \tgt{\hat{p}}} \leq 1 =
    x_{\src{p_1}, \tgt{p_1}} + 0 \leq 1
  \end{align*}
  as well as constraints \eqref{eq:num_proj_const}:
  \begin{multline*}
    \sum\limits_{t \in T} x_{\src{p}, t} \leq n_{proj} \implies                             \\
    \sum\limits_{t \in T} x*_{\src{p}, t} =
    \sum\limits_{t \in T \setminus \{ \tgt{p} \}} x_{\src{p}, t} + x^*_{\src{p}, \tgt{p}} =
    \sum\limits_{t \in T \setminus \{ \tgt{p} \}} x_{\src{p}, t} + 0 \leq n_{proj}
  \end{multline*}
  Therefore such a solution \( x^* \) is optimal as well as original solution \( x \), but
  target candidates have indices in the range \( [m, M] \).

  When it comes to the second option where a target candidate \( \tgt{\hat{p}} \) overlaps with some
  candidates from the shrunk set in a way that there is some aligned word between a source entity \( \src{p} \) and these target candidate, i.e.
  \( \left[ (i_{\tgt{\hat{p}}} < m) \lor (M < j_{\tgt{\hat{p}} < m}) \right] \land
  \left[ (i_{\tgt{\hat{p}}} \leq M) \lor m \leq j_{\tgt{\hat{p}}} \right] \).
  By the corollary \ref{col:shrink_cand} such a candidate \( \tgt{p} = ( \max(i_{\tgt{\hat{p}}}, m), \min(M, j_{\tgt{\hat{p}}}) ) \) will have
  higher aligned-based matching score.

  Let's consider a solution \( x^* \) that equals
  to a solution \( x \) everywhere except for target candidates \( \tgt{\hat{p}}  \) and \(  \tgt{p} \),
  i. e. \( x^*_{\src{p}, \tgt{\hat{p}}} = 0, x^*_{\src{p}, \tgt{p}} = 1 \). This solution also satisfies all constraints
  of the ILP problem. For constraints \eqref{eq:num_proj_const} the proof is the following:
  \begin{align*}
    & \sum\limits_{t \in T} x_{\src{p}, t} =
    \sum\limits_{t \in T \setminus \{ \tgt{p}, \tgt{\hat{p}} \}} x_{\src{p}, t} + x_{\src{p}, \tgt{p}} + x_{\src{p}, \tgt{\hat{p}}}  =     \\
    & \sum\limits_{t \in T \setminus \{ \tgt{p}, \tgt{\hat{p}} \}} x_{\src{p}, t} + 0 + 1
    \leq n_{proj} \implies                                                                                                                 \\
    & \sum\limits_{t \in T} x*_{\src{p}, t} =
    \sum\limits_{t \in T \setminus \{ \tgt{p}, \tgt{\hat{p}} \}} x_{\src{p}, t} + x^*_{\src{p}, \tgt{p}} + x^*_{\src{p}, \tgt{\hat{p}}}  = \\
    & \sum\limits_{t \in T \setminus \{ \tgt{p}, \tgt{\hat{p}} \}} x_{\src{p}, t} + 1 + 0 \leq n_{proj}
  \end{align*}
  And for non-overlapping constraints, since \( \tgt{p} \) is a substring of \( \tgt{\hat{p}} \) by a
  construction we have:
  \[
    \forall t \in T \quad t \cap \tgt{p} \neq \emptyset \implies t \cap \tgt{\hat{p}} \neq \emptyset
  \]
  It gets us that all constraints that should be hold for \( \tgt{p} \) should also be satisfied for \( \tgt{\hat{p}} \):
  \begin{equation} \label{eq:overlap_substring}
    \left\{ (\src{p_1}, \tgt{p_1}) \Big| (\src{p_1}, \src{p}, \tgt{p_1}, \tgt{p}) \in \hat{\Pi}(S, T) \right\} \subset
    \left\{ (\src{p_1}, \tgt{p_1}) \Big| (\src{p_1}, \src{p}, \tgt{p_1}, \tgt{\hat{p}}) \in \hat{\Pi}(S, T) \right\}
  \end{equation}
  Suppose that constraints \eqref{eq:non_overlap_const} are violated for the solution \( x^* \), then
  \begin{align*}
    & \exists \src{p_1} \in S, \tgt{p_1} \in T \Big| (\src{p_1}, \src{p}, \tgt{p_1}, \tgt{p}) \in \hat{\Pi}(S, T) \\
    & x^*_{\src{p_1}, \tgt{p_1}} + x^*_{\src{p}, \tgt{p}} =
    x_{\src{p_1}, \tgt{p_1}} + 1 > 1 \implies                                                                      \\
    & x_{\src{p_1}, \tgt{p_1}} = 1
    \implies                                                                                                       \\
    & x_{\src{p_1}, \tgt{p_1}} + x_{\src{p}, \tgt{\hat{p}}} = 1 + 1 = 2 > 1
  \end{align*}
  But because of \eqref{eq:overlap_substring}
  \( (\src{p_1}, \src{p}, \tgt{p_1}, \tgt{\hat{p}} ) \in \hat{\Pi}(S, T) \),
  therefore it contradicts that the solution \( x \) was feasible.
  Hence constraints \eqref{eq:non_overlap_const} are satisfied by the solution \( x^* \).

  And in the same time the solution \( x^* \) has a higher objective value that implies that the original solution \( x \) was not optimal.
  This contradicts our assumption that there is no such an optimal solution that ensures requirements of the theorem.
\end{proof}
Note that the theorem holds only for constraints \eqref{eq:num_proj_const} with a form of \( < \) or
\( \leq \) just because the ILP problem can lost feasibility if we remove projections with zero scores that
exist only to ensure these constraints and has no application sense.

In comparison to the heuristic algorithm that uses word-to-word alignment the formulation of the optimization
problem with matching scores \eqref{eq:align_cost} has an important advantage: we can evaluate the confidence of
every particular projection, since we have a score of it, whereas heuristics provide us only with labelling of the target
sentence. In the same time the prediction of this formulation and heuristics are not always the same, e.g.
the heuristic algorithm merges every continuous ranges of aligned words separated but at most some fixed number of non-aligned words,
but for the ILP formulation it will be so only for some specific cases.

Also it worth to mention that for the alignment-based score there is no fixed upper bound, since
the maximum value of the denominator is a product of length of a source entity and a target candidate.
It is possible to overcome it if we exchange one of the summation operation in the denominator from the
equation \eqref{eq:align_cost} with a logical OR. Nevertheless, during experiments we will use the
original form of the alignment-based matching score.

\subsection{NER model-based score}

An other way to evaluate matching scores between source entities and target candidates is to
use a multilingual NER model. The idea is the following: a model for every word of the target candidate
predicts probability distribution over a set of classes that a word has a specific label. In order to compute
a matching score that a source entity with a class \( l \) should be projected onto a target candidate,
we compute an average probability of the class \( l \) over all words of the candidate.

Let \( L \) be a set of classes. Since a NER model predicts output in the IOB format, i.e. the first word of a predicted
entity with class \( l \in L \) has a label \textit{B-l} and other words of this entity have labels \textit{I-l},
the output of a model can be represented as a matrix \( p_{m, o} \) where \( m \in \mathbb{N} \) is an index of a word and
\( o \in \{ 1, 2|L| + 1 \} \) is an index of a label (assume that the label \( O \), i. e. a word doesn't belong to any class,
has an index \( 2|L| + 1 \)).

Then let's introduce mappings \( B[l]: L \rightarrow \{ 1, \dots, 2|L| \} \) and
\( I[l]: L \rightarrow \{ 1, \dots, 2|L| \} \) that returns index of B and I labels respectively
for a class \( l \in L \) in a probability matrix.

So the NER model-based matching score is the following:
\begin{equation} \label{eq:ner_cost}
  c_{\src{p}, \tgt{p}}^{ner} = \alpha^{(j_{\tgt{p}} - i_{\tgt{p}}) - 1}
  \frac{
    p_{i_{\tgt{p}}, B[l_{\src{p}}]} +
    \sum\limits_{k = i_{\tgt{p}} + 1}^{j_{\tgt{p}}} p_{k, I[l_{\src{p}}]}
  }
  {j_\tgt{p} - i_{\tgt{p}}}
\end{equation}
where \( l_{\src{p}} \in L \) is a class of the source entity \( \src{p} \in S \) and
\( \alpha > 0 \in \mathbb{R} \) is a length-scaling constant.

The factor \( \alpha \) is necessary to align matching scores with NER model predictions.
We naturally expect from the matching cost that if a target candidate is a substring of a predicted
by a NER model entity, then its score should be lower that a score of the target candidate that
equal to the predicted entity. For example, let the set of classes contains only class \textit{PER}: \( L = \{ PER \}\),
set of source entities consists of only one entity \( \src{p} \) with a class \textit{PER},
set of target candidates be \( T =  \{ \tgt{p_1} = (1, 1), \tgt{p_2} = (1, 2) \} \) and a probability matrix predicted by the model is:
\[
  \begin{blockarray}{cccc}
    & \text{\textit{B-PER}} & \text{\textit{I-PER}} & \text{\textit{O}} \\
    \begin{block}{c(ccc)}
      1 & 0.9 & 0.1 & 0 \\
      2 & 0.2 & 0.8 & 0 \\
      3 & 0   & 0   & 1 \\
    \end{block}
  \end{blockarray}
\]
By taking maximum over rows it turns out that the entity predicted by a model has a class \textit{PER} and
consists of words with indices 1 and 2. But the NER model-based matching score without scaling factor
\( \alpha \) gives the following results:
\[
  c_{\src{p}, \tgt{p_1}}^{ner} = 0.9 \qquad c_{\src{p}, \tgt{p_2}}^{ner} = 0.85
\]
And in this case the ILP problem will prefer projecting source entity onto a substring of the entity predicted
by the model. Therefore, length-scaling constant \( \alpha \) can not be omitted.

If \( \alpha \) is a crucial part of the score then it make sense to evaluate the range of its values.
For this let's notice the following fact.
\begin{lemma} \label{lemma:maxprob_greater_than_avg}
  Let \( p_{i}, i \in K \) be a probability distribution over a finite set \( K \) with cardinality \( k \).
  Then
  \[
    \max\limits_{i \in K} p_i \geq \frac{1}{k}
  \]
\end{lemma}
\begin{proof}
  Suppose that it is wrong, i.e.
  \[
    \max\limits_{i \in K} p_i < \frac{1}{k}
  \]
  Then let's sum up all probabilities:
  \[
    \sum\limits_{i \in K} p_i \leq \sum\limits_{i \in K} \max\limits_{j \in K} p_j =
    k \cdot \max\limits_{j \in K} p_j < 1
  \]
  But the sum of all probabilities of the distribution should be equal to \( 1 \), therefore our assumption was
  wrong that proofs the lemma.
\end{proof}
Having this useful fact let's finally make some estimation of the constant \( \alpha \).
\begin{theorem}
  Consider a source entity \( \src{p} \in S \) and two target candidates \( \tgt{p_1} = (i, j+1), \tgt{p_2} = (i, j) \).
  Suppose the predictions \( p \) of a NER model satisfy the following constraints:
  \[
    B[l_{\src{p}}] = \argmax\limits_{l \in \{ 1, \dots, 2|L| + 1 \}} p_{i, l} \qquad
    I[l_{\src{p}}] = \argmax\limits_{l \in \{ 1, \dots, 2|L| + 1 \}} p_{k, l} \; \forall k \in \{ i + 1, \dots ,j +1 \}
  \]
  i.e. all these candidates are parts of the predicted by the model entity.
  Let's denote \( M = \frac{1}{2|L| + 1} \).
  Then:
  \[
    \alpha > 1 + \frac{1 - M}{1 + M} \implies c_{\src{p}, \tgt{p_2}}^{ner} > c_{\src{p}, \tgt{p_1}}^{ner}
  \]
\end{theorem}
\begin{proof} Let's write down a desired inequality
  \begin{align*}
    & c_{\src{p}, \tgt{p_2}}^{ner} > c_{\src{p}, \tgt{p_1}}^{ner} \Leftrightarrow \\
    & \alpha^{n}
    \frac{
      p_{i, B[l_{\src{p}}]} +
      \sum\limits_{k = i + 1}^{j + 1} p_{k, I[l_{\src{p}}]}
    }
    {n + 1} >
    \alpha^{n - 1}
    \frac{
      p_{i, B[l_{\src{p}}]} +
      \sum\limits_{k = i + 1}^{j} p_{k, I[l_{\src{p}}]}
    }
    {n}
    \Leftrightarrow                                                                \\
    & \alpha >
    \frac
    {(n+1) \cdot \left( p_{i, B[l_{\src{p}}]} +
    \sum\limits_{k = i + 1}^{j} p_{k, I[l_{\src{p}}]} \right)}
    {
      n \cdot \left( p_{i, B[l_{\src{p}}]} +
    \sum\limits_{k = i + 1}^{j + 1} p_{k, I[l_{\src{p}}]} \right)}
    =
    \frac
    {(n+1) \cdot \left( p_{i, B[l_{\src{p}}]} +
    \sum\limits_{k = i + 1}^{j} p_{k, I[l_{\src{p}}]} \right)}
    {
      n \cdot \left( p_{i, B[l_{\src{p}}]} +
    \sum\limits_{k = i + 1}^{j} p_{k, I[l_{\src{p}}]} + p_{j+1, I[l_{\src{p}}]} \right)}
  \end{align*}
  In order to simplify derivation we will introduce aliases for some quantities:
  \[
    S = p_{i, B[l_{\src{p}}]}  + \sum\limits_{k = i + 1}^{j} p_{k, I[l_{\src{p}}]} \qquad
    s = p_{j+1, I[l_{\src{p}}]}
  \]
  Note, that by the lemma \ref{lemma:maxprob_greater_than_avg} the following
  inequalities holds:
  \begin{equation} \label{eq:ner_sum_ineq}
    M \leq s \leq 1 \qquad nM \leq S \leq n
  \end{equation}
  Then we have:
  \begin{equation*}
    \alpha > \frac{(n + 1) S}{n (S + s)} =
    \frac{nS + S + ns - ns}{nS + ns} =
    1 + \frac{S - ns}{n(S + s)}
  \end{equation*}
  Let's compute derivatives of this expression.
  \begin{equation*}
    \left( 1 + \frac{S - ns}{n(S + s)} \right)_s^{'} =
    \frac{-n^2S - nS}{n^2 (S + s)^2} < 0
    \qquad
    \left( 1 + \frac{S - ns}{n(S + s)} \right)_S^{'}
    = \frac{nS + n^2s}{n^2 (S + s)^2} > 0
  \end{equation*}
  Hence the maximum of the function is located at the point
  \( s = M \) and \( S = n \), then we can take \( \alpha \) such that
  \begin{equation*}
    \alpha > 1 + \frac{1 - M}{1 + M}
    \stackrel{n \geq 1}{\geq}
    1 + \frac{n - nM}{n(n + M)}
    \stackrel{\max}{\geq}
    1 + \frac{S - ns}{n(S + s)}
  \end{equation*}
\end{proof}
As it is shown in the proof it is just an upper estimation of such an alpha for the worst possible case,
in practice there is a sense to try even smaller \( \alpha \), but then it doesn't guarantee that the
desired property are satisfied.

The main difference between a plain model transfer and XLNER pipeline with the ILP projection problem that uses NER model-based
matching cost is that during latter we project labels from source entities. Therefore some, possibly wrong, predictions of the
NER model could be ignored since in the solution of the ILP problem there is no source entity that projects
to this target entity.

One of the problem when it comes to using some model is the fact that models tend to be overconfident in their
predictions, i.e. outputs high probabilities for every, sometimes even wrong, prediction. And since the
proposed score directly uses these probabilities it can affects it. In order to overcome this issue the
calibration of the model should be applied. The simplest approach to do it is to scale output's logits of the
model by some temperature. Nevertheless, in further chapter we will use the NER model without any calibration.

The NER model-based matching score \eqref{eq:ner_cost} also has another property that is worth to mention.
Since the computation of the score involves only the class of the source entity, scores will be equal
for all source entities that have the same class. And from the point of view of the ILP problem
\eqref{eq:ilp} it can lead to a solution where a source entity is projected onto a semantically wrong candidate, but
nevertheless a correct class will be assigned to the target candidate.

\subsection{Translation-based score}

\begin{equation} \label{eq:nmt_cost}
  c_{\src{p}, \tgt{p}}^{nmt}
\end{equation}

\subsection{Fused score}
The performance of alignment, NER and translation models that were used for computation of
scores above varies depending on the language, domain, set of classes, etc.
Every matching score has their advantages and drawbacks, but the natural way to minimize fails is
to fuse scores together. Since all scores are real numbers one of the approaches to do it is to compute a
weighted sum of scores of all types:
\begin{equation} \label{eq:fused_cost}
  c_{\src{p}, \tgt{p}}^{fused} =
  \lambda_{align} c_{\src{p}, \tgt{p}}^{align} +
  \lambda_{ner} c_{\src{p}, \tgt{p}}^{ner} +
  \lambda_{nmt} c_{\src{p}, \tgt{p}}^{nmt}
\end{equation}
where \( \lambda_{align} \geq, \lambda_{ner} \geq 0, \lambda_{nmt} \geq 0 \in R\).

Moreover, the fused score can extend possibilities to handle issues of basic scores.
A NER model that is used in the NER-based cost \eqref{eq:ner_cost} sometimes can properly
predict spans of entities in the target sentence, but confuse classes.
We assume that labels are predicted correctly during computation of the NER model-based cost because
we have no way to correct them. But when the cost consists of different basic score, we can leverage a
NER model-based score only evaluate how likely a given candidate can form a target entity of any class and
determine label by using other types of scores. The modified \eqref{eq:ner_cost} that implements it
is the following:
\begin{equation} \label{eq:ner_cost_wo_classes}
  c_{\src{p}, \tgt{p}}^{ner} = \alpha^{(j_{\tgt{p}} - i_{\tgt{p}}) - 1}
  \max\limits_{l \in L}
  \frac{
    p_{i_{\tgt{p}}, B[l]} +
    \sum\limits_{k = i_{\tgt{p}} + 1}^{j_{\tgt{p}}} p_{k, I[l]}
  }
  {j_\tgt{p} - i_{\tgt{p}}}
\end{equation}

\section{Analysis of the ILP problem}
Despite the fact that the proposed ILP problem is just a particular case
of well-studied the general binary programming problem it doesn't imply that
it inherits all properties of it. For example, maximum bipartite matching problem
can be formulated as an instance of the ILP problem, but still can be solved in a
polynomial time, whereas general case of ILP is NP-hard. That motivates a deeper analysis
of the proposed formulation.

\subsection{Complexity}
Constraints, especially \eqref{eq:non_overlap_const}, play a crucial role in a complexity
of the proposed problem. In the case if all candidates are nor overlapping and constraints
\eqref{eq:num_proj_const} has a form of equalities with \( n_{proj} = 1 \) the problem reduces to
an instance of the weighted bipartite matching problem, that is in complexity class \( P \). But let's study
more general case.

For the sake of convenience we will use the first form \eqref{eq:objective}--\eqref{eq:binary_prog}
of the ILP problem in this section, but since it equivalent to the \eqref{eq:ilp} it doesn't affect
the results of the analysis.

The usual way to prove the complexity of some problem is to reduce other problem with known complexity to
an instance of the studied one. Thus let's consider the maximum independent set problem \cite{pemmaraju2003computational}.

\begin{figure}[ht]
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style = {draw, circle}]
      \node[fill=green!10] (1) {1};
      \node[right=of 1] (2) {2};
      \node[below right=of 2, fill=green!10] (3) {3};
      \node[below=of 1, fill=green!10] (4) {4};
      \node[left=of 4] (5) {5};
      \node[below right=of 4] (6) {6};

      \graph{
        (1) -- (2) -- (4) -- (6) -- (5) -- (4),
        (6) -- (2) -- (3)
      };
    \end{tikzpicture}
    \caption{Maximum independent set problem (MaxIS). Vertices that form an optimal solution are colored in \textbf{\textcolor{green!50}{green}}}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[
        node distance=0.1,
        realstate/.style = {draw, circle, font=\scriptsize},
        dummystate/.style = {draw, circle, minimum width=4, inner sep=6, fill=blue!10}
      ]
      \node[dummystate] (src_1) {};
      \node[right=of src_1, xshift=10, dummystate] (src_2) {};
      \node[right=of src_2, xshift=10, dummystate] (src_3) {};

      \node[below=of src_1, fill=green!10, xshift=-50, yshift=-1.75cm, realstate] (1) {1};
      \node[right=of 1, realstate] (2) {2};
      \node[right=of 2, fill=green!10, realstate] (3) {3};
      \node[right=of 3, fill=green!10, realstate] (4) {4};
      \node[right=of 4, realstate] (5) {5};
      \node[right=of 5, realstate] (6) {6};
      \node[right=of 6, dummystate](dummy_1) {};
      \node[right=of dummy_1] (dots) {\dots};
      \node[right=of dots, dummystate] (dummy_2) {};

      \node[above=of src_2, text=gray, font=\scriptsize] {Source nodes};
      \node[below=of 5, text=gray, font=\scriptsize] {Target nodes};

      \draw[green] (src_1) -- (1);
      \draw (src_1) -- node[above left, xshift=3, rotate=60, black, font=\footnotesize]{c=1} (2);
      \draw (src_1) -- (3);
      \draw (src_1) -- (4);
      \draw (src_1) -- (5);
      \draw (src_1) -- (6);
      \draw[gray!40] (src_1) -- (dummy_1);
      \draw[gray!40] (src_1) -- (dummy_2);
      \draw (src_2) -- (2);
      \draw[green] (src_2) -- (3);
      \draw (src_2) -- (4);
      \draw (src_2) -- (5);
      \draw (src_2) -- (6);
      \draw[gray!40] (src_2) -- (dummy_1);
      \draw[gray!40] (src_2) -- (dummy_2);
      \draw (src_3) -- (2);
      \draw (src_3) -- (3);
      \draw[green] (src_3) -- (4);
      \draw (src_3) -- (5);
      \draw (src_3) -- (6);
      \draw[gray!40] (src_3) -- (dummy_1);
      \draw[gray!40] (src_3) -- node[above right, black, font=\footnotesize]{c=0} (dummy_2);
    \end{tikzpicture}
    \caption{Reduction to the proposed ILP problem. Dummy nodes are colored in \textcolor{blue!60}{blue}}
  \end{subfigure}
  \caption{The diagram of the general idea of the reduction MaxIS problem to the proposed ILP problem}
  \label{fig:maxis_reduction}
\end{figure}

The maximum independent set problem is a problem of finding a subset of vertices of some undirected
graph with maximum cardinality, such that it doesn't contain any vertices connected by edges of the graph.
An example of the maximum independent set problem is depicted on the figure \ref{fig:maxis_reduction} (a).
The formal definition as an ILP problem is given by \ref{def:maxis}.
\begin{definition}[MaxIS] \label{def:maxis}
  Let \( G=(V, E), V \neq \emptyset \) be an undirected graph, then a maximum independent set problem for the graph \( G \) is the following:
  \begin{align*}
    & \max \sum\limits_{v \in V} x_v                               \\
    & x_u + x_v \leq 1               \qquad \forall \{u, v\} \in E \\
    & x_v \in \{0, 1\}
  \end{align*}
\end{definition}

The MaxIS problem is an NP-hard \cite{pemmaraju2003computational} problem since VertexCover can be reduced to an instance of
this problem. But what is even more interesting for us is that this problem looks very similar to the
our projection ILP problem, except the fact, that set of constraint \eqref{eq:num_proj_const} are
omitted and costs are equal to \( 1 \). Moreover, the relation whether two edges of the graph are adjacent
is not transitive as well as the relation of overlapping \ref{def:overlapping}! That's why we will try to reduce
the maximum independent set problem to an instance of the proposed ILP problem.

The idea of the reduction is straightforward. For every vertex \( v \in V \) of the MaxIS problem
we will create one distinct target candidate \( \tgt{p_v} \) that corresponds to this vertex.
The set of all such target candidate will be called \( T_V \). The ILP problem are formulated as a projection
from source entities, in order to represent MaxIS as this ILP problem let's create dummy source nodes. The number of source nodes
is determined in such a way that it will be easy to satisfy constraints~\eqref{eq:num_proj_const}.
The proposed number of dummy source nodes for any type of constraints \eqref{eq:num_proj_const} are given
in the table \ref{tab:dummy_nodes_num} (a). The matching cost between any dummy source node and
target candidate from the set \( T_V \) will make equal to \( 1 \). And finally let's link solution of the
ILP problem to the solutions of the MaxIS problem:
\[
  x_v = 1 \Leftrightarrow \exists \src{p} \in S \Big| x_{\src{p}, \tgt{p_v}} = 1
\]
i.e. the vertex \( v \in V \) belongs to the maximum independent set if and only if there is a dummy
source node that are projected to the target candidate \( \tgt{p_v} \) that correspond to this vertex.

\begin{table}[h]
  \begin{subtable}[t]{0.5\linewidth}
    \centering
    \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{|c|c|c|}
      \hline
      \(<\)                                 & \( \leq \), \( = \)               & \(>\), \( \geq \) \\
      \hline
      \( \ceil{\frac{|V|}{n_{proj} - 1}} \) & \( \ceil{\frac{|V|}{n_{proj}}} \) & 1                 \\
      \hline
    \end{tabular}
    \caption{Source nodes}
    \label{tab:src_dummy_nodes_num}
  \end{subtable}
  \begin{subtable}[t]{0.5\linewidth}
    \centering
    \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{|c|c|c|c|}
      \hline
      \(<\), \(\leq\) & \( = \)                                    & \(>\)          & \( \geq \)         \\
      \hline
      \( 0 \)         & \( \ceil{\frac{|V|}{n_{proj}}} n_{proj} \) & \( n_{proj} \) & \( n_{proj} + 1 \) \\
      \hline
    \end{tabular}
    \caption{Target nodes}
    \label{tab:tgt_dummy_nodes_num}
  \end{subtable}
  \caption{Number of dummy nodes for every type of constraints \eqref{eq:num_proj_const}}
  \label{tab:dummy_nodes_num}
\end{table}

We will consider that target nodes \( t_u, t_v \in T_{V} \) that correspond to graph's nodes \( u, v \in V \) overlap
if and only if there is an edge in the graph \( G \) between these nodes or \( u = v \).
\begin{equation} \label{eq:overlap_reduction}
  t_u \cap t_v \neq \emptyset \Leftrightarrow
  \left( \{ u, v \} \in E \right) \lor  \left(u = v \right)
\end{equation}
Such a definition of overlapping allow us to match all properties, i.e. reflexivity, symmetricity and non transitivity,
of overlapping relation on word ranges in the initial definition \ref{def:overlapping}. The fact that we consider
overlapping nodes that correspond to the same vertex simply means that it is impossible to
add the same vertex in the independent set two times which is a property of the any set.

Hence the resulting ILP problem induced by the MaxIS problem is the following:
\begin{equation} \label{eq:reduction_without_nproj}
  \begin{aligned}
    & \max\limits_x \sum\limits_{(\src{p}, \tgt{p}) \in S \times T} c_{\src{p}, \tgt{p}} x_{\src{p}, \tgt{p}}                                             \\
    & \text{subject to}                                                                                                                                   \\
    & x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1
    & \forall (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T)                                                                            \\
    & x_{\src{p}, \tgt{p}} \in \{ 0, 1 \}                                                                     & \forall (\src{p}, \tgt{p}) \in S \times T
  \end{aligned}
\end{equation}
where set \( \hat{PI} \) defined based on the introduced above overlapping relation.

The problem \eqref{eq:reduction_without_nproj} is exactly the same as the ILP problem \eqref{eq:objective}--\eqref{eq:binary_prog}
except the fact that constraints \eqref{eq:num_proj_const} are omitted. The reason is that we can not be sure
that them can be satisfied for type of constraints with \( =, >, \geq \) inequalities, since it is possible to just run out of nodes.
In order to overcome this problem, let's create dummy target nodes. The proposed minimum number of such nodes are given
in the table \ref{tab:dummy_nodes_num} (b). Let's denote set of all such nodes as \( T_{dummy} \) and then the set of all target candidates
will be \( T = T_V \cup T_{dummy} \). We will consider that dummy target nodes don't overlap with any node expect itself:
\[
  \forall \tgt{p_1} \in T_{dummy}, \; \tgt{p} \in T \quad
  \tgt{p_1} \cap \tgt{p_2} \neq \emptyset \Leftrightarrow \tgt{p_1} = \tgt{p_2}
\]
Target dummy nodes are created only to satisfy constraints \eqref{eq:num_proj_const}, so all matching scores
between any dummy source node and any dummy target node set to be equal \( 0 \).

Since we have a one to one correspondence between vertices from the MaxIS problem and target
candidates from the set \( T_V \), consider the following set that fully determine the solution of the
MaxIS problem:
\[
  T^*_{x} = \{ t \in T_V | \exists s \in S, x_{s, t} = 1 \}
\]
Let's \( x_1, x_2 \) be a two feasible solutions of problem \eqref{eq:reduction_without_nproj}, then we will say that
they are in relation~\( \sim \) if their corresponding solutions of the MaxIS problem are equal:
\[
  x_1 \sim x_2 \Leftrightarrow T^{*}_{x_1} = T^{*}_{x_2}
\]
Since this relation defined by an equality of sets it is reflexive, symmetric and transitive and therefore
equivalence relation.

Then quotient set \( \quot{X}{\sim} \) will consist of all equivalence classes of feasible solutions
that differ only in matching of dummy nodes. Every such equivalence class corresponds to one feasible solution of
the maximum independent set problem. Let's notice that the objective function value is equal for all
elements within any equivalence class.
\begin{lemma}
  Let X be a set of feasible solution of the problem \eqref{eq:reduction_without_nproj}. Then
  for any equivalence class in the quotient set \( \quot{X}{\sim} \) the objective
  value is the same for every solution within this equivalence class.
\end{lemma}
\begin{proof}
  Consider an equivalence class \( [x] \) of some feasible solution \( x \).
  Let's notice that non zero scores have only elements from the set \( T^*_x \),
  therefore the objective determines by this set of projected target candidates:
  \begin{multline*}
    \sum\limits_{(\src{p}, \tgt{p}) \in S \times T} c_{\src{p}, \tgt{p}} x_{\src{p}, \tgt{p}} =
    \sum\limits_{(\src{p}, \tgt{p}) \in S \times T_V} c_{\src{p}, \tgt{p}} x_{\src{p}, \tgt{p}} = \\
    \sum\limits_{(\src{p}, \tgt{p}) \in S \times T^{*}_{x}} c_{\src{p}, \tgt{p}} x_{\src{p}, \tgt{p}}
  \end{multline*}
  Because of non-overlapping constraints and the fact that we consider that target candidate overlaps with
  itself only one source entity can be projected onto every target candidate from the set \( T_x^* \):
  \[
    \forall \tgt{p} \in T^*_x, \; \exists! \src{p} \in S \Big| x_{\src{p}, \tgt{p}} = 1
  \]
  Then the objective value equals to the cardinality of the set \( T^*_x \) that are equal for
  any solution without the same equivalence class by the definition of the relation \( ~ \).
  \[
    \sum\limits_{(\src{p}, \tgt{p}) \in S \times T} c_{\src{p}, \tgt{p}} x_{\src{p}, \tgt{p}} =
    \sum\limits_{(\src{p}, \tgt{p}) \in S \times T^{*}_{x}} c_{\src{p}, \tgt{p}} x_{\src{p}, \tgt{p}} =
    |T_x^*|
  \]
\end{proof}
\begin{corollary} \label{col:maxis_and_ilp_objective_equal}
  The objective function value of the ILP problem \eqref{eq:reduction_without_nproj} and the maximum independent
  set problem are equal.
\end{corollary}
\begin{proof}
  By construction every target candidate from the set \( T^*_x \) corresponds to one only one vertex
  from the MaxIS problem. Therefore the objective value is equal to the cardinality of the set \(  T^*_x \).
  In the proof of the lemma we showed that the objective value of the ILP problem~\eqref{eq:reduction_without_nproj}
  equals to \( | T^*_x| \) as well.
\end{proof}

Dummy source nodes exists only to be able to formulate the MaxIS problem as the desired ILP projection problem.
Hence we can swap a dummy source node for a target candidate it was projected from.
\begin{lemma} \label{lemma:swap_source_reduction}
  Let \( x \) be a feasible solution of the problem \eqref{eq:reduction_without_nproj} such that
  some dummy source node \( \src{p} \in S \) is projected onto \( \tgt{p} \in T \),
  i.e. \( x_{\src{p}, \tgt{p}} = 1 \). Then for any other source node
  \( \src{\hat{p}} \in S, \src{p} \neq \src{\hat{p}} \) a solution \( x^{*} \), such that
  \begin{align*}
    & x^*_{\src{\hat{p}}, \tgt{p}} = 1  \qquad
    x^*_{\src{p}, \tgt{p}} = 0                                                                                            \\
    & \forall (s, t) \in S \times T \setminus \{ (\src{p}, \tgt{p}), (\src{\hat{p}}, \tgt{p}) \} \quad x^*_{s,t} = x_{s,t}
  \end{align*}
  is also feasible and in the same equivalence class as \( x \).
\end{lemma}
\begin{proof}
  Suppose that the solution \( x^* \) is not feasible, it means that:
  \[
    \exists s \in S, t \in T \Big| (s, \src{\hat{p}}, t, \tgt{p}) \in \hat{\Pi}(S, T)
    \quad x^*_{s, t} = 1
  \]
  Let's notice that \( s \neq \src{p} \), since \( x^*_{\src{p}, \tgt{p}} = 0 \).
  But then by construction:
  \[
    x_{\src{p}, \tgt{p}} + x_{s, t} = 1 + x^*_{s, t} = 1 + 1 > 2
  \]
  It contradicts with the fact that \( x \) is a feasible solution, therefore \( x^* \) should be
  feasible.

  And since by construction sets \( T^*_x \) and \( T^*_{x^*} \) are equal solutions are in the same equivalence
  class and by the corollary \ref{col:maxis_and_ilp_objective_equal} the objective values for these solutions are also equal.
\end{proof}

\begin{corollary} \label{col:bound_num_proj_reduction}
  For any feasible solution of the problem \eqref{eq:reduction_without_nproj} there is a feasible solution  \( x^* \)
  such that
  \[
    \sum\limits_{\tgt{p} \in T} x^*_{\src{p}, \tgt{p}} \leq \frac{|V|}{|S|}
    \qquad \forall \src{p} \in S
  \]
  and \( T_{x^*}^* =  T^*_x \).
\end{corollary}
\begin{proof}
  Assume the solution \( \hat{x} \) such that
  \begin{align*}
    & \forall \src{p} \in S \;\; \forall \tgt{p} \in T_V \qquad \hat{x}_{\src{p}, \tgt{p}} = x_{\src{p}, \tgt{p}} \\
    & \forall \src{p} \in S \;\; \forall \tgt{p} \in T_{dummy} \qquad \hat{x}_{\src{p}, \tgt{p}} = 0
  \end{align*}
  By construction this solution has the same objective value since all score that for dummy target nodes are equal to \( 0 \),
  the same set \( T^*_{\hat{x}} = T^*_x \)
  and also it doesn't violate the non-overlapping constraints:
  \begin{align*}
    & \forall (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T) \Big| \tgt{p_2} \in T_{V} \quad
    \hat{x}_{\src{p_1}, \tgt{p_1}} + \hat{x}_{\src{p_2}, \tgt{p_2}} =
    x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1                                                      \\
    & \forall (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T) \Big| \tgt{p_2} \in T_{dummy} \quad
    \hat{x}_{\src{p_1}, \tgt{p_1}} + 0 \leq
    x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1
  \end{align*}

  Because of non-overlapping constraints \eqref{eq:non_overlap_const} at most one source node can
  be projected on every target node, therefore:
  \[
    \sum\limits_{\src{p} \in S} \sum\limits_{\tgt{p} \in T} \hat{x}_{\src{p}, \tgt{p}} \leq |T_V| \leq |V|
  \]
  Suppose there exists such a \( \src{p} \in S \) that
  \[
    \sum\limits_{\tgt{p} \in T} \hat{x}_{\src{p}, \tgt{p}} > \frac{|V|}{|S|}
  \]
  Then there is another source entity \( \src{\hat{p}} \) for which we have
  \[
    \sum\limits_{\tgt{p} \in T} \hat{x}_{\src{\hat{p}}, \tgt{p}} < \frac{|V|}{|S|}
  \]
  By the lemma \ref{lemma:swap_source_reduction} there exists such an optimal solution \( x^* \)
  where
  \begin{align*}
    & \sum\limits_{\tgt{p} \in T} x^*_{\src{p}, \tgt{p}} = \sum\limits_{\tgt{p} \in T} \hat{x}_{\src{p}, \tgt{p}} - 1             \\
    & \sum\limits_{\tgt{p} \in T} x^*_{\src{\hat{p}}, \tgt{p}} = 1 + \sum\limits_{\tgt{p} \in T} \hat{x}_{\src{\hat{p}}, \tgt{p}} \\
  \end{align*}
  and \( T^*_x = T^*_{\hat{x}} = T^*_{x^*}  \). If this solution doesn't satisfy the desired property repeat the procedure.
\end{proof}

Let's note that all feasible solutions of the MaxIS will correspond to feasible solutions of the
problem \ref{eq:reduction_without_nproj} and vice versa.

\begin{lemma} \label{lemma:maxis_f_implies_ilp}
  Suppose \( x_v \) is a feasible solution of the maximum independent set problem, then there is a
  corresponding solution of the problem \eqref{eq:reduction_without_nproj}.
\end{lemma}
\begin{proof}
  Suppose some a non feasible solution \( x \) that corresponds to the feasible solution of the
  MaxIS problem. Then we have:
  \[
    \exists (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T) \Big|
    x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} > 1
  \]
  But by construction of the set \( \hat{\Pi}(S, T) \) and the overlapping relation (dummy target nodes
  are overlapping only with itself) if \( \tgt{p_1} \neq \tgt{p_2} \) then:
  \begin{multline*}
    \left\{ (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \Big| (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T) \right\} = \\
    \left\{ (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \Big| (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T_V) \right\}
  \end{multline*}
  But since the solution \( x_v \) is feasible constraints with  \( \tgt{p_1} \neq \tgt{p_2} \) can not be violated:
  \begin{align*}
    & \forall \{ u, w \} \in E \quad x_u + x_w \leq 1 \stackrel{\eqref{eq:overlap_reduction}}{\implies}                  \\
    & \forall \tgt{p_1} \in T^*_x \; \nexists \tgt{p_2} \in T^*_x \Big| \tgt{p_1} \cap \tgt{p_2} \neq \emptyset \implies \\
    & \forall \src{p_1}, \src{p_2} \in S \quad x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1
  \end{align*}
  Therefore \( \tgt{p_1} = \tgt{p_2} \) and constraints are violated because there exists
  at least two source nodes \( \src{p_1}, \src{p_2} \in S \) that are projected onto the same target nodes \( \tgt{p} \in T \):
  \[
    x_{\src{p_1}, \tgt{p}} = 1 \qquad x_{\src{p_2}, \tgt{p}} = 1
  \]

  But then consider a solution \( x^* \) such that we remove one of the projection and make this constraint :
  \begin{align*}
    & \forall (s,t) \in S \times T \setminus \left\{ (\src{p_2}, \tgt{p}) \right\}
    \quad x^*_{s,t} = x_{s,t} \\
    & x^*_{\src{p_2}, \tgt{p}} = 0
  \end{align*}
  Note that \( T^*_{x^*} = T^*_x  \) because there is a projection to the target node \( \tgt{p} \)
  since \( \)  \( x^*_{\src{p_1}, \tgt{p}} = 1 \).
  If the solution \( x^* \) is infeasible - repeat procedure, otherwise
  this solution is feasible and corresponds to the same solution \( x_v \) of the MaxIS problem.
\end{proof}

\begin{lemma} \label{lemma:ilp_f_implies_maxis}
  Suppose \( x \) is a feasible solution of the problem \eqref{eq:reduction_without_nproj}, then
  the corresponding solution of the maximum independent set problem is feasible as well.
\end{lemma}
\begin{proof} By construction the non-overlapping constraints of the problem \eqref{eq:reduction_without_nproj}
  imply constraints of the MaxIS problem:
  \begin{align*}
    & \forall (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T) \quad
    x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1 \implies                                                     \\
    & \forall (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T) \Big| \tgt{p_1}, \tgt{p_2} \in T_V \quad
    x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} \leq 1 \implies                                                     \\
    & \forall \tgt{p_1}, \tgt{p_2} \in T_V, \tgt{p_1} \neq \tgt{p_2}, \tgt{p_1} \cap \tgt{p_2} \neq \emptyset
    \quad \nexists \src{p_1}, \src{p_2} \in S \Big|
    x_{\src{p_1}, \tgt{p_1}} + x_{\src{p_2}, \tgt{p_2}} > 1 \implies                                                     \\
    & \nexists \tgt{p_1}, \tgt{p_2} \in T_V, \tgt{p_1} \neq \tgt{p_2}, \tgt{p_1} \cap \tgt{p_2} \neq \emptyset \Big|
    \tgt{p_1} \in T^*_x, \tgt{p_2} \in T^*_x \implies                                                                    \\
    & \forall \{ v_{\tgt{p_1}}, v_{\tgt{p_2}} \} \in E \quad x_{v_{\tgt{p_1}}} + x_{v_{\tgt{p_2}}}\leq 1
  \end{align*}
\end{proof}

And finally we can prove the complexity of the proposed ILP problem by reduction the MaxIS problem
to an instance of \eqref{eq:objective}--\eqref{eq:binary_prog}.
\begin{theorem}[\( MaxIS \leq_P ILP_{proj} \)]
  The proposed ILP problem \eqref{eq:objective}--\eqref{eq:binary_prog} is NP-hard
\end{theorem}
\begin{proof}
  Assume the reduction of the MaxIS problem to an instance of the projection ILP problem described above.
  Since number of nodes in the ILP problem is linear on the number of vertices in the MaxIS problem and
  having a solution of the ILP problem it scales at most quadratically on number of vertices time to
  determine the solution of the MaxIS problem the reduction takes a polynomial time.
  The only thing that it is required to check is whether the projected optimal solution of the ILP problem
  will be always optimal solution of the MaxIS and vice versa.

  By the corollary \ref{col:maxis_and_ilp_objective_equal} the objective values of two problems are equal and therefore
  feasible solution with highest objective value will have a highest objective value in the counterpart problem.
  And by lemmas \ref{lemma:maxis_f_implies_ilp} and \ref{lemma:ilp_f_implies_maxis} the non-overlapping constraint
  won't make any optimal solution infeasible for their counterpart.

  But besides non-overlapping constraint the problem \eqref{eq:objective}--\eqref{eq:binary_prog} also has
  constraints \eqref{eq:num_proj_const}. Let's check whether they won't make any optimal solution of the MaxIS
  problem infeasible in their ILP formulation and there are no new optimal solutions.

  Consider a set \( X \) of all feasible solutions of the problem \eqref{eq:reduction_without_nproj} and
  set \( Y \) of all feasible solutions of the problem \eqref{eq:objective}--\eqref{eq:binary_prog}. Since
  the latter problem is constrained version of the problem \eqref{eq:reduction_without_nproj} \( Y \subset X \), but then:
  \[
    Y \subset X \implies \forall y \in Y \quad \exists x \in X \Big| \quad [y] \subset [x],
  \]
  i.e. there won't be no new equivalence classes and therefore no new feasible solutions of the MaxIS problem.
  But it turns out that for any equivalence class there is a feasible solution of the full problem that belongs to this class:
  \[
    \forall x \in X \quad \exists y \in Y \Big| \quad y \in [x]
  \]
  Let's prove it. For this we need to analyze constraints \eqref{eq:num_proj_const} for every type of inequalities.

  \textit{The case of \( < \).} The number of dummy source nodes is given in the table \ref{tab:dummy_nodes_num} (a) and equal to
  \( |S| = \ceil{\frac{|V|}{n_{proj} - 1}} \)
  By the corollary \ref{col:bound_num_proj_reduction} there exists such a solution \( x^* \) from the same
  equivalence class that satisfies constraints:
  \[
    \forall \src{p} \in S \quad
    \sum\limits_{\tgt{p} \in T} x^*_{\src{p}, \tgt{p}} \leq \frac{|V|}{|S|} =
    \frac{|V|}{ \ceil{\frac{|V|}{n_{proj} - 1}}} \leq \frac{|V|}{\frac{|V|}{n_{proj} - 1}} =
    n_{proj} - 1 < n_{proj}.
  \]

  \textit{The case of \( \leq \).} By the same corollary \ref{col:bound_num_proj_reduction} as in the previous case
  we obtain that there is a solution \( x^* \) that satisfies constraints:
  \[
    \forall \src{p} \in S \quad
    \sum\limits_{\tgt{p} \in T} x^*_{\src{p}, \tgt{p}} \leq \frac{|V|}{|S|} =
    \frac{|V|}{ \ceil{\frac{|V|}{n_{proj}}}} \leq \frac{|V|}{\frac{|V|}{n_{proj}}} =
    n_{proj} .
  \]

  \textit{The case of \( =\).} By the corollary \ref{col:bound_num_proj_reduction}
  for any solution \( x \) from the equivalence class \( [x] \) there is
  a feasible solution \( x^* \in [x] \) such that:
  \begin{equation} \label{eq:red_thm_n_proj}
    \forall \src{p} \in S \quad
    \sum\limits_{\tgt{p} \in T} x^*_{\src{p}, \tgt{p}} \leq \frac{|V|}{|S|} =
    \frac{|V|}{ \ceil{\frac{|V|}{n_{proj}}}} \leq \frac{|V|}{\frac{|V|}{n_{proj}}} =
    n_{proj} .
  \end{equation}
  If constraints \eqref{eq:num_proj_const} are violated then there is such a source node \( \src{\hat{p}} \)
  for which the sum is a strong inequality:
  \begin{equation} \label{eq:red_thm_src_that_violate_constr}
    \exists \src{\hat{p}} \in S \quad
    \sum\limits_{\tgt{p} \in T} x^*_{\src{p}, \tgt{p}} \leq \frac{|V|}{|S|} <
    n_{proj} .
  \end{equation}
  But then we can find such a dummy target node \( t_d \in T_{dummy} \) that:
  \[
    \exists t_d \in T_{dummy} \Big| \forall s \in S \quad x^*_{s,t} = 0
  \]
  It can be proved by contradiction to the property of the given by the corollary \ref{col:bound_num_proj_reduction}
  and that there exists such a \( \src{\hat{p}} \):
  \begin{align*}
    & \nexists t_d \in T_{dummy} \Big| \forall s \in S \quad x^*_{s,t} = 0 \implies                                                             \\
    & \forall t \in T_{dummy}, \; \exists s \in S \quad x^*_{s,t} = 1 \implies                                                                  \\
    & \ceil{\frac{|V|}{n_{proj}}} n_{proj} = |S| n_{proj}
    \stackrel{\eqref{eq:red_thm_n_proj}}{\geq}
    \sum\limits_{s \in S} \sum\limits_{t \in T} x^*_{s,t} \geq |T_{dummy}| = \ceil{\frac{|V|}{n_{proj}}} n_{proj} \implies                       \\
    & \sum\limits_{s \in S} \sum\limits_{t \in T} x^*_{s,t} = \ceil{\frac{|V|}{n_{proj}}} n_{proj} \geq |V|
    \implies                                                                                                                                     \\
    & \sum\limits_{s \in S \setminus \{ \src{\hat{p}} \}} \sum\limits_{t \in T} x^*_{s,t} = \ceil{\frac{|V|}{n_{proj}}} n_{proj} = |S| n_{proj}
    \stackrel{\eqref{eq:red_thm_src_that_violate_constr}}{\implies}                                                                              \\
    & \exists s \in S \quad
    \sum\limits_{\tgt{p} \in T} x^*_{\src{p}, \tgt{p}} > \frac{|V|}{|S|} >
    n_{proj}
  \end{align*}
  In short words always there are a "free" dummy target nodes that can be used to ensure equality constraints.
  It was expectable just because during construction we choose the number of target dummy nodes to make it possible.
  In order to make the solution closer to be feasible we just need to project source entity \( \src{\hat{p}} \) onto this target
  candidate \( t_d \). So, the modified solution \( \hat{x} \) is the following:
  \begin{align*}
    & \forall (s,t) \in S \times T \setminus \left\{ (\src{\hat{p}}, t_d) \right\} \hat{x}_{s,t} = x^*_{s,t} \\
    & \hat{x}_{\src{\hat{p}}, t_d} = 1
  \end{align*}
  By the definition of the set \( T^*_{\hat{x}} \) since we add only a projection to a dummy node
  it won't change the equivalence class of the solution, objective value but make the number of projection
  from the source entity \( \src{\hat{p}} \) higher:
  \[
    \sum\limits_{\tgt{p} \in T} \hat{x}_{\src{\hat{p}}, \tgt{p}} =
    \sum\limits_{\tgt{p} \in T} x^*_{\src{\hat{p}}, \tgt{p}} + 1
    \leq n_{proj}
  \]
  Since dummy target nodes overlaps only with itself we have that the constraints that potentially
  can be violated are defined by this set:
  \begin{multline*}
    \left\{ (\src{p_1}, \src{p_2}, t_d, \tgt{p_2}) \Big| (\src{p_1}, \src{p_2}, \tgt{p_1}, \tgt{p_2}) \in \hat{\Pi}(S, T) \right\} = \\
    \left\{ (\src{p_1}, \src{p_2}, t_d, t_d) \Big| \src{p_1}, \src{p_2} \in S, \src{p_1} \neq \src{p_2} \right\}
  \end{multline*}
  But because of the choice of \( t_d \) the constraints
  \eqref{eq:non_overlap_const} are not violated:
  \begin{align*}
    \forall \src{p_1}, \src{p_2} \in S \Big| \src{p_1} \neq \src{p_2}, \src{p_1} \neq \src{\hat{p}}, \src{p_2} \neq \src{\hat{p}} \quad
    & \hat{x}_{src{p_1}, t_d} + \hat{x}_{\src{p_2}, t_d} = x_{\src{p_1}, t_d} + x_{\src{p_2}, t_d} = 0 \\
    \forall \src{p_2} \in S \Big| \src{\hat{p}} \neq \src{p_2} \quad
    & \hat{x}_{\src{\hat{p}}, t_d} + \hat{x}_{\src{p_2}, t_d} = 1 + 0 = 1 \leq 1
  \end{align*}

  If still there are some source entity for which constrains \eqref{eq:num_proj_const} are violated - repeat the process.
  Repetition of the process is possible since the property \eqref{eq:red_thm_n_proj} still holds and
  we proved that there always exists a "free" dummy target node we can project this source node onto.

  \textit{The case of \( >, \geq \).} We can perform a similar derivation as for the case of \( = \) type of constraints, but
  let's note that \( \forall \src{p} \in S \):
  \begin{align*}
    & \sum\limits_{\tgt{p} \in T} x_{\src{p}, \tgt{p}} = n_{proj} \Leftrightarrow
    n_{proj} \geq \sum\limits_{\tgt{p} \in T} x_{\src{p}, \tgt{p}} \geq n_{proj}                                  \\
    & \bigvee\limits_{k=n_{proj} + 1}^{|T|} \sum\limits_{\tgt{p} \in T} x_{\src{p}, \tgt{p}} = k \Leftrightarrow
    \sum\limits_{\tgt{p} \in T} x^*_{\src{p}, \tgt{p}} > n_{proj} \qquad
  \end{align*}
  Therefore all solutions that are feasible for the case of equality constraints are also should be
  feasible for the case of \( \geq, > \) and vice versa, so we won't lose any equivalence class.

  So, we have shown that for every feasible solution of the problem \eqref{eq:reduction_without_nproj} there is
  a feasible solution of the full ILP problem \eqref{eq:objective}--\eqref{eq:binary_prog} that belongs to the
  same equivalence class. And every feasible solution of the full ILP problem is a feasible solution of the problem \eqref{eq:reduction_without_nproj}.
  Therefore the MaxIS problem can be solved as an instance of the problem \eqref{eq:objective}--\eqref{eq:binary_prog}
  and consequently the ILP problem is at least as hard as the maximum independent set problem which is NP-hard.
\end{proof}

\subsection{Approaches to compute the solution of the problem}
As any ILP problem the problem \eqref{eq:ilp} can be solved by branch and bound, cutting planes,
branch and cut methods, other exact algorithms. Nevertheless it can take a lot of time to find an optimal solution and
it can make such formulation inefficient from the application point of view where it is required to
solve instances of this problem for thousands sentences in a limited time.

It motivates the necessity of the approximate algorithm that can compute a solution that is not always
optimal, but don't violate crucial constraints. One of the approach for such an algorithm for the problem
\eqref{eq:ilp} is to iteratively assign \( 1 \) to a variable with the highest matching score and then remove
all target candidates that overlap with the projected candidate to enforce non-overlapping constraints.
Algorithm \ref{alg:ilp_greedy} is the variant of a such greedy algorithm.

\begin{algorithm}
  \caption{Approximate greedy algorithm for the proposed ILP problem} \label{alg:ilp_greedy}
  \KwData{instance of the ILP problem \eqref{eq:ilp}}
  \KwResult{\( x \) -- "solution" of the ILP problem}

  \( x \gets 0 \) \;
  \( P \gets 0 \) \Comment*[r]{number of projections by source entity}
  \While{\( \exists \src{p} \in S, \tgt{p} \in T \Big| c_{\src{p}, \tgt{p}} > 0 \)}{
    \( s, t \gets \argmax\limits_{\src{p} \in S, \tgt{p} \in T } c_{\src{p}, \tgt{p}} \) \;
    \( x_{s,t} \gets 1 \) \;
    \( P_{s} \gets P_{s} + 1 \) \;

    \ForAll(\tcp*[f]{remove all overlapping with \( t \) candidates}){\( \hat{t} \in T \Big| \hat{t} \cap t \neq \emptyset \)}{
      \( c_{s, \hat{t}} \gets 0 \) \;
    }

    \Comment{try to ensure constraints \eqref{eq:num_proj_const}}
    \If{constraints \eqref{eq:num_proj_const} is a type of \( =, \leq \)}{
      \If{\( P_s = n_{proj} \)}{
        \ForAll{\( \hat{t} \in T \)}{
          \( c_{s, \hat{t}} \gets 0 \) \;
        }
      }
    }
    \If{constraints \eqref{eq:num_proj_const} is a type of \( < \)}
    {
      \If{\( P_s = n_{proj} - 1 \)}{
        \ForAll{\( \hat{t} \in T \)}{
          \( c_{s, \hat{t}} \gets 0 \) \;
        }
      }
    }
  }
\end{algorithm}

From the steps of the algorithm \ref{alg:ilp_greedy} it implies that after removing all
overlapping target candidates by making their costs equal to zero the algorithm ensures the
constants \eqref{eq:num_proj_const} on a number of projected from every source entity candidates.
But only for the case where constraints have a form of \( <, \leq \).

In the case of \( =, >, \geq \) the algorithm will output a "solution" that violates constraint
\eqref{eq:num_proj_const} only in two cases. The first one is when there is a source entity for which
all target candidates, that are not overlapping with already projected ones, initially have zero matching cost. But then,
from the application point of view it doesn't make sense to project the source entity onto these candidates
since it is definitely not a counterpart of the source entity in the target sentence, otherwise it would not
have a matching cost equal to \( 0 \). The second option is the situation where target candidates
that could potentially be projected onto have been removed on the previous iterations of the algorithm.
And it is hard to handle this issue without backtracking and losing feasibility on
non-overlapping constraints \eqref{eq:non_overlap_const}.

Nevertheless, the algorithm \ref{alg:ilp_greedy} is linear on a number of variables and therefore can
solve the ILP problem significantly faster than the exact ILP solver in a general case.

Even in the case when the output of the greedy algorithm is a feasible solution it does not
necessary mean that the solution is optimal, i.e. there may exists a feasible solution with higher objective
value. This facts leads from the fact that the proposed ILP problem is NP-hard and therefore polynomial
algorithm can not solve it in a general case unless P=NP. But also it can be easily shown on an example.
Consider the ILP problem \eqref{eq:ilp} with \( n_{proj} = 2 \) and \( \leq \) type of constraints \eqref{eq:num_proj_const}.
Let \( T = \{ \tgt{p_1} = (1, 2), \tgt{p_2} = (2, 3), \tgt{p_3} = (3, 5) \}, S = \{ \src{p} \} \) and
the matching scores are the following:
\[
  c_{\src{p}, \tgt{p_1}} = 0.2 \qquad
  c_{\src{p}, \tgt{p_2}} = 0.3 \qquad
  c_{\src{p}, \tgt{p_3}} = 0.2
\]
The output of the algorithm \ref{alg:ilp_greedy} is the solution \( \src{p} \) with only one
non zero variable where \( x_{\src, \tgt{p_2}} = 1 \). It has an objective value \( 0.3 \).
While the optimal solution is the one where source entity is projected onto \( \tgt{p_1} \) and
\( \tgt{p_2} \) with the objective value equal to \( 0.4 \).
